---
title: "Datathong 2026 fall1"
author: "TIan Yuan"
date: "2026-10-03"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
library(shiny)
library(dplyr)
library(ggplot2)
library(quanteda)
library(Matrix)
library(shiny)
library(pROC)
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

```{r , echo=FALSE}
Sys.setenv(LANGUAGE = "en")
library(quanteda)
library(ggplot2)
data<-read.csv("January2015toNovember2025.csv",stringsAsFactors = FALSE)
names(data)
dim(data)
str(data)
colSums(is.na(data))
sum(is.na(data$NatureTitle))
sum(duplicated(data))
sort(colMeans(is.na(data)), decreasing = TRUE)
class(data$EventDate)
range(data$EventDate)
data$EventDate <- as.Date(
  data$EventDate,
  format = "%m/%d/%Y"
)
class(data$EventDate)
library(readxl)
library(dplyr)
naic<-read_xlsx("2022_NAICS_Structure.xlsx",skip=2)
names(naic)
naic$`2022 NAICS Code` <- as.character(naic$`2022 NAICS Code`)
data <- data |>
  mutate(
    Primary.NAICS = as.character(Primary.NAICS)
  ) |>
  left_join(
    naic,
    by = c("Primary.NAICS" = "2022 NAICS Code")
  )
names(data)

head(
  data |>
    select(Primary.NAICS, `2022 NAICS Title`)
)

sum(is.na(data$`2022 NAICS Title`))
data |>
  filter(is.na(`2022 NAICS Title`)) |>
  count(Primary.NAICS, sort = TRUE) |>
  head(30)
#calculate proportion of each year
data |>
  mutate(
    Year = format(EventDate, "%Y"),
    NAICS_matched = !is.na(`2022 NAICS Title`)
  ) |>
  group_by(Year) |>
  summarise(
    Total = n(),
    Unmatched = sum(!NAICS_matched),
    Unmatched_Percent = round(
      mean(!NAICS_matched) * 100,
      2
    )
  )
data <- data |>
  mutate(
    NAICS2 = substr(Primary.NAICS, 1, 2),

    Sector = case_when(
      NAICS2 == "11" ~ "Agriculture, Forestry, Fishing and Hunting",
      NAICS2 == "21" ~ "Mining, Quarrying, and Oil and Gas Extraction",
      NAICS2 == "22" ~ "Utilities",
      NAICS2 == "23" ~ "Construction",
      NAICS2 %in% c("31", "32", "33") ~ "Manufacturing",
      NAICS2 == "42" ~ "Wholesale Trade",
      NAICS2 %in% c("44", "45") ~ "Retail Trade",
      NAICS2 %in% c("48", "49") ~ "Transportation and Warehousing",
      NAICS2 == "51" ~ "Information",
      NAICS2 == "52" ~ "Finance and Insurance",
      NAICS2 == "53" ~ "Real Estate and Rental and Leasing",
      NAICS2 == "54" ~ "Professional, Scientific, and Technical Services",
      NAICS2 == "55" ~ "Management of Companies and Enterprises",
      NAICS2 == "56" ~ "Administrative and Support and Waste Management and Remediation Services",
      NAICS2 == "61" ~ "Educational Services",
      NAICS2 == "62" ~ "Health Care and Social Assistance",
      NAICS2 == "71" ~ "Arts, Entertainment, and Recreation",
      NAICS2 == "72" ~ "Accommodation and Food Services",
      NAICS2 == "81" ~ "Other Services (except Public Administration)",
      NAICS2 == "92" ~ "Public Administration",
      TRUE ~ "Unknown"
    )
  )
table(data$Sector)
sector_count <- data |>
  count(Sector)
ggplot(sector_count, aes(x = n, y = Sector)) +
  geom_col() +
  scale_fill_brewer(palette = "Blues") +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey70", linewidth = 0.2),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "sector_count",
    x = "n",
    y = "sectors"
  )
#injury by occupation
data |>
  group_by(Sector) |>
  summarise(
    Reports = n(),
    Hospitalized = sum(Hospitalized)
  )
#calculate hospi rt#
sector_summary <- data |>
  group_by(Sector) |>
  summarise(
    Reports = n(),
    Hospitalized = sum(Hospitalized)
  )
sector_summary <- mutate(
  sector_summary,
  Hospitalization_Rate = Hospitalized / Reports
)
ggplot(sector_summary, aes(x = Hospitalization_Rate, y = Sector)) +
  geom_col() +
  scale_fill_brewer(palette = "Blues") +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey70", linewidth = 0.2),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "hospitallrate ",
    x = "%",
    y = "sectors"
  )
count(data, EventTitle, sort = TRUE)
count(data, NatureTitle, sort = TRUE)
#idea
###################################################################################
sort(unique(data$NatureTitle))
data$Nature_Group <- ifelse(
  grepl("fracture", data$NatureTitle, ignore.case = TRUE), "Fracture",
  ifelse(
    grepl("amputation|avulsion|enucleation", data$NatureTitle, ignore.case = TRUE), "Amputation",
    ifelse(
      grepl("burn|corrosion", data$NatureTitle, ignore.case = TRUE), "Burn",
      ifelse(
        grepl("sprain|strain|tear", data$NatureTitle, ignore.case = TRUE), "Sprain/Strain/Tear",
        ifelse(
          grepl("cut|laceration|puncture|open wound", data$NatureTitle, ignore.case = TRUE), "Open Wound",ifelse(
            grepl("soreness|pain|hurt", data$NatureTitle, ignore.case = TRUE), "Pain/Soreness",

            ifelse(
              grepl("intracranial|concussion|brain", data$NatureTitle, ignore.case = TRUE), "Brain/Head Injury",

              ifelse(
                grepl("crush", data$NatureTitle, ignore.case = TRUE), "Crushing",

                ifelse(
                  grepl("internal injur", data$NatureTitle, ignore.case = TRUE), "Internal Injury",

                  ifelse(
                    grepl("heat exhaustion|heat and light|heat", data$NatureTitle, ignore.case = TRUE), "Heat-related",

                    ifelse(
                      grepl("poison", data$NatureTitle, ignore.case = TRUE), "Poisoning",

                      ifelse(
                        grepl("trauma", data$NatureTitle, ignore.case = TRUE), "Trauma - Unspecified",
                        ifelse(
                          grepl("Bruis|contusi", data$NatureTitle, ignore.case=TRUE), "Bruise",
                          ifelse(
                            grepl("disloca",data$NatureTitle, ignore.case=TRUE),"Dislocation",
                            ifelse(
                            grepl("gunsho",data$NatureTitle, ignore.case=TRUE),"Gun",
                            ifelse(
                            grepl("allerg",data$NatureTitle, ignore.case=TRUE),"allergy related",
                            ifelse(
                            grepl("soft tissue",data$NatureTitle, ignore.case=TRUE),"soft tissue",
                              
                            
                          
          "Other"
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
  )
)
)
)
)
)
)
table(data$Nature_Group)
other <- data[data$Nature_Group == "Other", ]
head(
  sort(table(other$NatureTitle), decreasing = TRUE),
  30
)
prop.table(table(data$Amputation, useNA = "ifany")) * 100
table(data$Amputation, useNA = "ifany")
table(data$Hospitalized, useNA = "ifany")
prop.table(table(data$Hospitalized, useNA = "ifany")) * 100
data <- data |>
  mutate(
    Hospitalized_binary = ifelse(Hospitalized >= 1, 1, 0)
  )
table(data$Loss.of.Eye, useNA = "ifany")

prop.table(
  table(data$Loss.of.Eye, useNA = "ifany")
) * 100
data <- data |>
  mutate(
    Amputation_binary = ifelse(Amputation >= 1, 1, 0),
    EyeLoss_binary = ifelse(Loss.of.Eye >= 1, 1, 0)
  )
```
```{r , echo=FALSE}

model_data <- data |>
  select(
    Amputation_binary,
    Sector,
    EventTitle,
    SourceTitle,
    State,
    EventDate
  ) |>
  na.omit()
str(data)

n_distinct(model_data$EventTitle)
n_distinct(model_data$SourceTitle)
n_distinct(model_data$Sector)
n_distinct(model_data$State)
data |>
  count(EventTitle) |>
  summarise(
    Total_Event_Types = n(),
    Min_Count = min(n),
    Median_Count = median(n),
    Max_Count = max(n)
  )
head(
  data |>
    select(Event, EventTitle),
  30
)
data <- data |>
  mutate(
    Event_char = as.character(Event),
    Event_Division = substr(Event_char, 1, 1)
  )
data |>
  count(Event_Division, sort = TRUE)
data <- data |>
  mutate(
    Event_Group = case_when(
      Event_Division == "1" ~ "Violence / Persons / Animals",
      Event_Division == "2" ~ "Transportation Incidents",
      Event_Division == "3" ~ "Fires and Explosions",
      Event_Division == "4" ~ "Falls, Slips, Trips",
      Event_Division == "5" ~ "Harmful Substances / Environments",
      Event_Division == "6" ~ "Contact with Objects / Equipment",
      Event_Division == "7" ~ "Overexertion / Bodily Reaction",
      Event_Division == "9" ~ "Nonclassifiable / Other",
      TRUE ~ "Unknown"
    ))
head(
  data |>
    select(Source, SourceTitle),
  30
)
data <- data |>
  mutate(
    Source_char = as.character(Source),
    Source_Division = substr(Source_char, 1, 1)
  )
data|>
  count(Source_Division)
data <- data |>
  mutate(
    Source_Group = case_when(
      Source_Division == "1" ~ "Chemicals and Chemical Products",
      Source_Division == "2" ~ "Containers, Furniture, and Fixtures",
      Source_Division == "3" ~ "Machinery",
      Source_Division == "4" ~ "Parts and Materials",
      Source_Division == "5" ~ "Persons, Plants, Animals, and Minerals",
      Source_Division == "6" ~ "Structures and Surfaces",
      Source_Division == "7" ~ "Tools, Instruments, and Equipment",
      Source_Division == "8" ~ "Vehicles",
      Source_Division == "9" ~ "Other Sources",
      TRUE ~ "Unknown"
    )
  )
```

```{r , echo=FALSE}
library(caret)
data <- data |>
  filter(!is.na(Amputation_binary))
set.seed(123)
index <- createDataPartition(
  model_data$Amputation_binary,
  p = 0.8,
  list = FALSE
)
dattrain <- data[index, ]
dattest <- data[-index, ]


prop.table(table(data$Amputation_binary))

prop.table(table(dattrain$Amputation_binary))

prop.table(table(dattest$Amputation_binary))
set.seed(123)
#folds/ cross validation
folds <- createFolds(
  dattrain$Amputation_binary,
  k = 5,
  list = TRUE
)
val_index <- folds[[1]]

cv_validation <- dattrain[val_index, ]
cv_training <- dattrain[-val_index, ]
#mod reg
mod <- glm(
  Amputation_binary ~ Sector + Event_Group + Source_Group + State,
  data = cv_training,
  family = binomial)
pred_prob <- predict(
  mod,
  newdata = cv_validation,
  type = "response"
)
head(pred_prob)
#thhdd 0.5
pred_class <- ifelse(
  pred_prob >= 0.5,
  1,
  0
)

head(pred_class)
#conf mat
confusionMatrix(
  factor(pred_class, levels = c(0, 1)),
  factor(cv_validation$Amputation_binary, levels = c(0, 1)),
  positive = "1"
)
#14102 good
#No Information Rate : 0.7325          
#P-Value [Acc > NIR] : <2e-16    
# Sensitivity : 0.6773        identify how many cases                      **highlight**        
##             Specificity : 0.8879          
##          Pos Pred Value : 0.6882          
##          Neg Pred Value : 0.8828          
##              Prevalence : 0.2675          
##          Detection Rate : 0.1812          
##    Detection Prevalence : 0.2633          
##       Balanced Accuracy : 0.7826          
cv_resul <- data.frame(
  Fold = 1:5,
  Accuracy = NA,
  Sensitivity = NA,
  Specificity = NA,
  Balanced_Accuracy = NA
)
for (i in 1:5) {

  val_index <- folds[[i]]

  cv_validation <- dattrain[val_index, ]
  cv_training <- dattrain[-val_index, ]

  model <- glm(
    Amputation_binary ~ Sector + Event_Group + Source_Group + State,
    data = cv_training,
    family = binomial
  )

  pred_prob <- predict(
    model,
    newdata = cv_validation,
    type = "response"
  )
  roc_obj <- roc(
    cv_validation$Amputation_binary,
    pred_prob
)

  cv_resul$AUC[i] <- as.numeric(auc(roc_obj))
  pred_class <- ifelse(pred_prob >= 0.5, 1, 0)

  cm <- confusionMatrix(
    factor(pred_class, levels = c(0, 1)),
    factor(cv_validation$Amputation_binary, levels = c(0, 1)),
    positive = "1"
  )

  cv_resul$Accuracy[i] <- cm$overall["Accuracy"]
  cv_resul$Sensitivity[i] <- cm$byClass["Sensitivity"]
  cv_resul$Specificity[i] <- cm$byClass["Specificity"]
  cv_resul$Balanced_Accuracy[i] <- cm$byClass["Balanced Accuracy"]
}
cm
#see folds perform
cv_resul
colMeans(cv_resul[, -1])
#stable quite
#wanna use:pred_prob
library(pROC)
roc_obj <- roc(
  cv_validation$Amputation_binary,
  pred_prob
)
auc(roc_obj)
#new col in cvresul
cv_resul$AUC <- NA
roc_obj <- roc(
  cv_validation$Amputation_binary,
  pred_prob
)

cv_resul$AUC[i] <- as.numeric(auc(roc_obj))
#notes:
##   Fold  Accuracy Sensitivity Specificity Balanced_Accuracy       AUC
## 1    1 0.8315839   0.6773198   0.8879317         0.7826258 0.9007618
## 2    2 0.8279380   0.6734244   0.8843893         0.7789069 0.8942793
## 3    3 0.8329501   0.6790881   0.8895701         0.7843291 0.8983809
## 4    4 0.8294610   0.6784897   0.8818716         0.7801807 0.8935418
## 5    5 0.8273381   0.6786441   0.8798372         0.7792406 0.8957046
##          Accuracy       Sensitivity       Specificity Balanced_Accuracy 
##         0.8298542         0.6773932         0.8847200         0.7810566 
##               AUC 
##         0.8965337
#stable***
#final mod(likely)                 check***
modfin<-glm(
  Amputation_binary ~ Sector + Event_Group + Source_Group + State,
  data = dattrain,
  family = binomial
)
test_prob <- predict(
  modfin,
  newdata = dattest,
  type = "response"
)

test_class <- ifelse(test_prob >= 0.5, 1, 0)
test_cm <- confusionMatrix(
  factor(test_class, levels = c(0, 1)),
  factor(dattest$Amputation_binary, levels = c(0, 1)),
  positive = "1"
)
test_roc <- roc(
  dattest$Amputation_binary,
  test_prob,
  quiet = TRUE
)
test_cm
test_cm
auc(test_roc)

```
```{r, echo=FALSE}
incident_data <- data |>
  filter(
    !is.na(Final.Narrative),
    nzchar(trimws(as.character(Final.Narrative)))
  ) |>
  mutate(
    CaseID = row_number(),
    Narrative = as.character(Final.Narrative)
  )
```

### Workplace Hazard Scenario

```{r, echo=FALSE}
selectInput(
  "sector",
  "Industry Sector:",
  choices = sort(unique(na.omit(data$Sector)))
)

selectInput(
  "event",
  "Event Type:",
  choices = sort(unique(na.omit(data$Event_Group)))
)

selectInput(
  "source",
  "Source Type:",
  choices = sort(unique(na.omit(data$Source_Group)))
)

selectInput(
  "state",
  "State:",
  choices = sort(unique(na.omit(data$State)))
)
```

```{r, echo=FALSE}
new_case <- reactive({
  data.frame(
    Sector = input$sector,
    Event_Group = input$event,
    Source_Group = input$source,
    State = input$state
  )
})

amputation_prob <- reactive({
  as.numeric(
    predict(
      modfin,
      newdata = new_case(),
      type = "response"
    )
  )
})

historical_frequency <- reactive({
  data |>
    filter(
      Sector == input$sector,
      Event_Group == input$event,
      Source_Group == input$source,
      State == input$state
    ) |>
    summarise(Reports = n())
})
```

### Scenario Indicators

```{r, echo=FALSE}
renderText({
  probability <- amputation_prob()

  if (!is.finite(probability)) {
    return("Predicted Amputation Likelihood: Unavailable")
  }

  paste0(
    "Predicted Amputation Likelihood: ",
    round(probability * 100, 1),
    "%"
  )
})
```

```{r, echo=FALSE}
renderText({
  paste0(
    "Historical Severe Injury Reports: ",
    historical_frequency()$Reports
  )
})
```

### Workplace Injury Priority Matrix

```{r, echo=FALSE}
renderPlot({

  frequency <- historical_frequency()$Reports
  probability <- amputation_prob() * 100

  validate(
    need(
      is.finite(probability),
      "Prediction is unavailable for this scenario."
    )
  )

  plot_data <- data.frame(
    Frequency = frequency,
    Amputation = probability
  )

  ggplot(
    plot_data,
    aes(x = Frequency, y = Amputation)
  ) +
    geom_point(
      size = 5,
      color = "red"
    ) +
    scale_y_continuous(
      limits = c(0, 100)
    ) +
    labs(
      title = "Workplace Injury Scenario",
      subtitle = "Historical reports and estimated amputation likelihood",
      x = "Historical Severe Injury Reports",
      y = "Predicted Amputation Likelihood (%)"
    ) +
    theme_minimal(base_size = 14)
})
```

### Top 5 Reported Incident Patterns

```{r, echo=FALSE}
renderPlot({

  top_patterns <- data |>
    filter(Sector == input$sector) |>
    count(
      Event_Group,
      name = "Reports",
      sort = TRUE
    ) |>
    slice_head(n = 5)

  validate(
    need(
      nrow(top_patterns) > 0,
      "No incidents found for this sector."
    )
  )

  ggplot(
    top_patterns,
    aes(
      x = Reports,
      y = reorder(Event_Group, Reports)
    )
  ) +
    geom_col(fill = "steelblue") +
    labs(
      title = "Top 5 Reported Incident Patterns",
      subtitle = paste("Industry:", input$sector),
      x = "Number of Severe Injury Reports",
      y = "Event Type"
    ) +
    theme_minimal(base_size = 12)
})
```

### Incident Case Finder

```{r, echo=FALSE}
incident_cases <- reactive({

  data |>
    filter(
      Sector == input$sector,
      Event_Group == input$event,
      Source_Group == input$source,
      State == input$state
    ) |>
    arrange(desc(EventDate)) |>
    select(
      EventDate,
      State,
      EventTitle,
      SourceTitle,
      Hospitalized,
      Amputation,
      Final.Narrative
    ) |>
    slice_head(n = 10) |>
    mutate(
      EventDate = as.character(EventDate)
    )
})
```

```{r, echo=FALSE}
renderTable({
  incident_cases()
}, striped = TRUE, bordered = TRUE)
```

## Incident Similarity Finder

Search historical OSHA severe injury reports using a written incident description.

### Prepare Text Search

```{r, echo=FALSE}
incident_tokens <- tokens(
  incident_data$Narrative,
  remove_punct = TRUE,
  remove_numbers = TRUE
)

incident_tokens <- tokens_tolower(
  incident_tokens
)

incident_tokens <- tokens_remove(
  incident_tokens,
  pattern = stopwords("en")
)

incident_tokens <- tokens_wordstem(
  incident_tokens,
  language = "english"
)

incident_dfm <- dfm(incident_tokens)

incident_tfidf <- dfm_tfidf(
  incident_dfm,
  scheme_tf = "count",
  scheme_df = "inverse"
)

case_matrix <- as(
  incident_tfidf,
  "dgCMatrix"
)

case_norms <- sqrt(
  Matrix::rowSums(case_matrix ^ 2)
)

idf <- log10(
  ndoc(incident_dfm) /
    pmax(docfreq(incident_dfm), 1)
)
```

### Similarity Search Function

```{r, echo=FALSE}
search_incidents <- function(
  query,
  top_n = 5,
  event_filter = "All"
) {

  empty_results <- incident_data[0, ] |>
    mutate(
      Similarity = numeric(0),
      EventDate = as.character(EventDate)
    ) |>
    select(
      Similarity,
      EventDate,
      State,
      EventTitle,
      SourceTitle,
      Amputation,
      Narrative
    )

  query_tokens <- tokens(
    query,
    remove_punct = TRUE,
    remove_numbers = TRUE
  )

  query_tokens <- tokens_tolower(
    query_tokens
  )

  query_tokens <- tokens_remove(
    query_tokens,
    pattern = stopwords("en")
  )

  query_tokens <- tokens_wordstem(
    query_tokens,
    language = "english"
  )

  query_dfm <- dfm(query_tokens)

  query_dfm <- dfm_match(
    query_dfm,
    features = featnames(incident_dfm)
  )

  query_matrix <- as(
    query_dfm,
    "dgCMatrix"
  )

  query_matrix <- query_matrix %*%
    Matrix::Diagonal(x = idf)

  query_norm <- sqrt(
    sum(query_matrix ^ 2)
  )

  if (!is.finite(query_norm) || query_norm == 0) {
    return(empty_results)
  }

  similarity <- as.numeric(
    case_matrix %*% Matrix::t(query_matrix)
  ) / (case_norms * query_norm)

  similarity[!is.finite(similarity)] <- 0

  # Apply event filter before ranking
  if (event_filter != "All") {

    excluded <- is.na(incident_data$Event_Group) |
      incident_data$Event_Group != event_filter

    similarity[excluded] <- 0
  }

  best <- order(
    similarity,
    decreasing = TRUE
  )

  best <- head(
    best[similarity[best] > 0],
    top_n
  )

  if (length(best) == 0) {
    return(empty_results)
  }

  incident_data[best, ] |>
    mutate(
      Similarity = round(similarity[best], 3),
      EventDate = as.character(EventDate)
    ) |>
    select(
      Similarity,
      EventDate,
      State,
      EventTitle,
      SourceTitle,
      Amputation,
      Narrative
    )
}
```

### Search Historical Incidents

```{r, echo=FALSE}
textInput(
  "incident_query",
  "Describe an incident:",
  placeholder = "A worker injured his hand while cleaning machinery"
)

selectInput(
  "incident_event",
  "Filter by Event Type:",
  choices = c(
    "All",
    sort(unique(na.omit(incident_data$Event_Group)))
  ),
  selected = "All"
)

actionButton(
  "search_button",
  "Find Similar Incidents"
)

search_results <- eventReactive(
  input$search_button,
  {

    req(nzchar(trimws(input$incident_query)))

    search_incidents(
      query = input$incident_query,
      top_n = 5,
      event_filter = input$incident_event
    )
  }
)

renderTable({

  results <- search_results()

  validate(
    need(
      nrow(results) > 0,
      "No matching incidents found. Try a different description or event type."
    )
  )

  results

}, striped = TRUE, bordered = TRUE)
```
